\bf What is the Best Way to Display Two Time Series to Read Association?

### 2016-02-09 paper/paper.Rnw
supp material

To allow for a greater control of distributions of the data in the lineups, we use simulated data. The data must contain autocorrelation for there to be meaningful information in the time dimension. To fulfil this requirement, each plot has a standardised pair of AR(1) models defined by:

@@

With $t \in \left\{12, 24, 48, 96\right\}$ and $e_{t}$ being generated with from a bivariate normal distribution:

@@

The variation in the covariance parameter allows for control in the correlation in the pairs of data. The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$. True data pairs are generated with $\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$. Pilot testing found that the rate of selecting the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly selecting the true plot, for pairs with correlation below $\rho = 0.3$. Further to this, relationships weaker than this point are often of little interest to analysts. Hence we found it unnecessary to test a more full range of correlations. Simulating data does not produce correlations exactly equal to $\rho$, so simulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$. The null pairs often had correlation generated spuriously through the simulation process. This was particularly problematic for generating small sample data, but setting $\beta = min(0.5,100t)$ reduced spurious correlation to manageable levels.
For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications. To counter this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component. This filter creates a trend by penalising both errors in smoothness and in fit. The trend of the series, $\tau_{t}$ is defined by the equation:

@@

Setting $\lambda = 1$ produced a smoothing that is more consistent with real world time series applications and did not introduce excessive spurious null correlation. However, for $t = \{12, 24\}$, any form of smoothing was unfeasible without introducing extreme null correlation, hence smoothing was only used for $t = \{48, 96\}$. Each generated lineup was produced as both a scatter plot and an overlaid line graph.

@@

\begin{itemize}
  \item Factor 1: Plot design, levels = scatter plot, overlaid time series.
  \item Factor 2: Sample size (t), levels = $\{12, 24, 48, 96\}$.
  \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$.
\end{itemize}

>>>>

For there to be meaningful information in the time dimension, the data must have autocorrelation. To fulfil this requirement, each plot has a pair of standardised AR(1) models generated by:

@@

With $t \in \left\{12, 24, 48, 96\right\}$ and $e_{t}$ generated from a bivariate normal distribution:

@@

The variation in the covariance parameter allows for control in the correlation in the pairs of data. The null data has no underlying structure, so uncorrelated pairs are produced by setting $\rho = 0$. True data pairs are generated with $\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$. Pilot testing found that the rate of selecting the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly selecting the true plot, for pairs with correlation below $\rho = 0.3$. Further to this, relationships weaker than this point are often of little interest to analysts. Hence we found it unnecessary to test a more full range of correlations. Simulating data does not produce correlations exactly equal to $\rho$, so simulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$. The null pairs often had correlation generated spuriously through the simulation process, which was particularly problematic for small sample data. Setting $\beta = min(0.5,100t)$ reduced spurious correlation to manageable levels.
For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications. To reduce this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component. This filter creates a trend by penalising both errors in smoothness and in fit. The trend of each series $y_{t}$, $\tau_{t}$ is defined by the equation:

@@

Setting $\lambda = 1$ produced a smoothing that is more consistent with real world time series applications and did not introduce excessive spurious null correlation. However, for $t = \{12, 24\}$, smoothing becomes unfeasible without introducing extreme null correlation, often with an absolute value greater than 0.8. Smoothing was only used for $t = \{48, 96\}$. Each real data set was placed with a collection of 19 null plots and produced as both a scatter plot and an overlaid line graph.

@@

\begin{itemize}
  \item Factor 1: Plot design, levels = scatter plot, overlaid time series.
  \item Factor 2: Sample size, t, levels = $\{12, 24, 48, 96\}$.
  \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$.
\end{itemize}
